---
title: "Homework 4 - A.Sauk"
author: "Alexandra Sauk"
date: "25/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2 Variables and Hypotheses

I hypothesize that class affected passenger survival as different classes were located on different levels of ship leading to different distances from the life boats.

I hypothesize that age affected passenger survival, either in ability to reach the life boats or preferrential placement on the life boats.

I hypothesize that gender affected passenger survival, either in ability to reach the life boats or preferrential placement on the life boats.

I hypothesize that the number of parents and/or children on board affected passenger survival as those with relatives on board may have prioritized survival of relatives. 

I hypothesize that cabin number affected passenger survival as certain cabins would have been located closer to the crash site and the midline of the ship where it cracked. 

## 3 Mosaic plots of the relationship of categorical predictor variables with survival

```{r}
my.data<-read.csv("titanic.csv")
head(my.data)

library("vcd")
mosaic(survived~pclass, data=my.data)
mosaic(survived~Gender, data=my.data)
```

## 3 logi.hist.plot of the relationship of continuous predictor variables with survival

```{r}
library("popbio")

my.data.parch<-na.omit(data.frame("relatives"=my.data$parch, "survived"=my.data$survived))
logi.hist.plot(my.data.parch$relatives, my.data.parch$survived, boxp=FALSE, type="hist", col="grey", xlab="Number of parents and/or children")

```



